---
title: "lab_12"
author: "Mitchell Hang"
date: "2023-05-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## You will need

-   tidytext and our usual libraries

## Load libraries and establish settings

**Task** Create a codeblock and load appropriate packages and settings for this lab.

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse.
library(tidytext)
library(tidyverse)
library(dplyr)
library(janitor)
library(lubridate)
library(rvest)
```

## Questions

**Q1.** You've been assigned to report a story about the leading reasons that Maryland attorneys get sanctioned by the state for misconduct. The state [publishes lists of sanctions](https://www.courts.state.md.us/attygrievance/sanctions) that contain a short text description about the situation. Load the CSV file in the data folder containing records from fiscal year 2011 onwards. Make a list of unique words from the text column, then following the example in the pre_lab, remove common "stop words" from that list and create a list of the top 10 words containing the percentage of occurrences each word represents. What's the leading word in that answer and, broadly, what do you think the top 10 words describe?

```{r}
#Load the data
md_attorney_sanctions <- read_csv("data/md_attorney_sanctions.csv")
```

```{r}
#Create list of unique words

unique_words <- md_attorney_sanctions |>
  select(text) |>
  unnest_tokens(word,text)

```

```{r}
#Load stop words
data("stop_words")

stop_words <- stop_words |>
  add_row(word = 'in') |> 
  add_row(word = 'disbarred') |>
  add_row(word = "an") |> 
  add_row(word = "and") |> 
  add_row(word = "to") |>
  add_row(word = "be") |>
  add_row(word = "for") |>
  add_row(word = 'maryland') |> 
  add_row(word = 'attorney') |> 
  add_row(word = 'md') |> 
  add_row(word = 'but') |> 
  add_row(word = "of") |> 
  add_row(word = "the") |>
  add_row(word = 'attorney') |>
  add_row(word = 'trust') |>
  add_row(word = 'client') |>
  add_row(word = 'respondent')
```

```{r}
#Remove stop words from unique words list
unique_words_no_stop_words <- unique_words |>
  anti_join(stop_words)

unique_words_no_stop_words |>
  group_by(word) |>
  tally(sort=TRUE) |>
  mutate(percent = (n/sum(n))*100) |>
  top_n(10)
```

**A1. We find from this brief analysis that a lot of the words relate to common terms to be expected such as client, respondent, trust. However, it was interested that the top word was failing. After adding more customized stop words, the top words are even more specific, including account, funds, conduct, consent, representation and reprimand.**

------------------------------------------------------------------------

**Q2.** Let's move beyond single words to phrases. Make a list of the top 10 three-word phrases, called trigrams, based on the example from the pre_lab (you'll need to modify the example code to do this). What's the top trigram and how often does it appear? What does that phrase mean in legal terms?

```{r}
# Check the task 12 on Pre-lab 12 to see if you can use that code
md_attorney_sanctions |>
  unnest_tokens(trigram, text, token = "ngrams", n = 3) |>
  separate(trigram, c("word1", "word2", "word3"), sep = " ") |>
  filter(!word1 %in% stop_words$word) |>
  filter(!word2 %in% stop_words$word) |>
  filter(!word3 %in% stop_words$word) |>
  mutate(trigram = paste(word1, word2, word3, sep=" ")) |>
  group_by(trigram) |>
  tally(sort=TRUE) |>
  mutate(percent = (n/sum(n))*100) |>
  top_n(10)
```

**A2. The top trigram is conduct involving dishonesty and appears 155 times. This term means wrongful acts committed by an attorney such as theft, embezzlement of money, or the wrongful taking or conversion of money, property, or other things of value, regardless of where the conduct occurred.**

------------------------------------------------------------------------

**Q3.** Let's drop back down to more traditional text analysis - take the top trigram from Q2 and write code to see how many times it occurs in the text column in each fiscal year. What do you think the answer produced by your code suggests? What else could you do to try and clarify the most important reasons attorneys get sanctioned?

```{r}
md_attorney_sanctions |>
  filter(str_detect(text, 'attorney trust account')) |>
  group_by(fiscal_year) |>
  tally(sort = TRUE)

```

**A3.** My top answer was conduct involving dishonesty due to filtering for stop words. This appeared 33 times. This suggests that attorneys got sanctioned for misconduct during the pandemic due to hte highest incidents occurring between 2019 and 2021. Another way to clarify this is to look at other trigrams and see if there are other higher ranking forms of misconduct.
