---
title: "lab_08"
author: "derek willis"
date: "2023-04-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and establish settings

**Task** Create a codeblock and load appropriate packages and settings for this lab. We'll be making some charts, working with dates and retrieving Census data.

```{r}
options(scipen=999)

library(tidyverse)
library(tidycensus)
library(ggplot2)
```

## Load and modify data

**Task** Create a codeblock and load the following data from the data folder:

-   Prince George's County 911 Overdose Calls

You will need to create columns for the date, week and month based on the existing `datetime` column.

```{r}
pg_call <- read_csv("data/prince_georges_2022_overdoses.csv")

pg_call <- pg_call |> 
  mutate(
    date = date(datetime), 
    month = month(datetime), 
    week = week(datetime)
    )
```

## Questions

**Q1.** Which month saw the greatest percentage of total calls? Create a dataframe that calculates the percentage of all calls that each month's calls represents. Make a bar chart from that dataframe. Your bar chart must have:

-   A clear title that states the main idea/finding
-   Good labels for the x & y axis and a caption for the source, which is Prince George's County EMS.
-   Readable bars - the values shouldn't be overlapping

Compare the results here to those from the pre_lab_08 bar chart - are there major differences in the months with the highest and lowest figures? Describe that below.

**A1. The three major differences include that the x axis in the pre-lab was the names of all of the months and the lab is just the number that corresponds to the month. Also, the scale is completely different in the pre-lab versus the lab for the y-axis. In the lab we based it off of the percentages and the pre-lab used the actual total number of calls. Lastly, the titles are different. The pre-lab uses a basic title and the lab uses a title from the actual findings.**

```{r}
pg_call_month <- pg_call |>
  group_by(month) |>
  summarize(total_calls = n()) |>
  mutate(pct = total_calls/1397*100) |>
  arrange(desc(total_calls))
  
pg_call_month |>
  ggplot() +
  geom_bar(aes(x=reorder(x=month,pct), weight=pct)) +
  coord_flip() +
  labs(
    title="December has one in 10 overdose calls in PG county",
    x = "Month",
    y = "Percentage of Calls",
    caption = "source: Prince George's County"
    )
```

**Q2.** Let's visualize this data on a weekly basis using a line chart. As in Q1, generate a dataframe with the total number of calls for each week, and then create a line chart to show the distribution of calls over time. Your line chart must have:

-   A clear title that states the main idea/finding
-   Good labels for the x & y axis and a caption for the source, which is Prince George's County EMS.
-   Readable labels

Describe the pattern of results; you may want to look at the data to dig into particular outliers.

**A2. The results of this line chart show a general scale to which the calls fall within but there is a major drop-off at the end that makes me question the significance of that value after 50 weeks have passed. Also, there is a major spike of calls in the center of the data.**

```{r}
pg_call_week <- pg_call |>
  group_by(week) |>
  summarise(total_calls = n())

pg_call_week |>
  ggplot() +
  geom_line(aes(x=week, y=total_calls)) +
  labs(
    title="Overdose Calls per week with a major spike in the middle of the year and a drop off in the ending weeks",
    x = "Weeks",
    y = "Percentage of Calls",
    caption = "source: Prince George's County"
    )
```

**Q3.** A scatterplot is a type of chart that helps us see relationships between two variables. One variable goes on the x axis, the other on the y axis. For each row/observation in our data, a scatterplot puts a circle (or a "point") where the two variables intersect on a grid.

Statisticians use scatterplots to show graphically whether one variable is correlated -- related, in a statistical sense -- with another variable. A classic example is the [relationship between ice cream sales and temperature](https://www.mathsisfun.com/data/scatter-xy-plots.html). The scatterplot below -- press play to load the image -- shows that relationship, that an increase in temperature is associated with an increase in ice cream sales. When it's 12C, sales are 200 dollars, and when it's hotter, 25C, sales are 600 dollars.

```{r}
knitr::include_graphics("https://www.mathsisfun.com/data/images/scatter-ice-cream1.svg")
```

We're going to use a scatterplot a little differently, to get a visual sense of two key variables:

Our question is: **does the median income in a zip code have any relationship to the number of overdose 911 calls in that zip code?**

To answer this question, do the following:

1.  Generate a dataframe with the number of 911 calls for each zip code.
2.  Get data from the Census Bureau showing median household income for Maryland zip codes.
3.  Join those two dataframes on their zip code columns, starting with the 911 calls dataframe.
4.  Make a scatterplot showing the total calls and median income. I didn't show you how to do this, so look it up! Googling "ggplot scatterplot" is a good start.
5.  Give it an appropriate title, source, and x and y axis titles.
6.  Add a label for each point that shows the zip code using geom_text() - see some examples of its use at <https://ggplot2.tidyverse.org/reference/geom_text.html#ref-examples>. Try to make the names as easy to read as possible by avoiding overlap.
7.  In the answer space below, describe what you see and answer the questions posed above. In a general sense, what do you think this means? Feel free to consider the actual raw values: how would you report out the main point(s) of this chart?

**A3. The scatterplot clearly shows that median income has a relationship with the number of overdose calls in each zip code across Maryland. The graph indicates that as median income decreases, the number of overdose calls increases. The zip code indicates a level of median income, and where it falls with the number of overdose calls. I would report the main points of this chart by looking at the zip code 20748, which has a low median income and a high number of overdose calls. I would compare this to the zip code 20623, which has a high median income and low number of overdose calls.**

```{r}

#1 Generate a dataframe with the number of 911 calls for each zip code.
calls_per_zipcode <- pg_call |>
  group_by(zipcode) |>
  summarise(total_calls = n())

colnames(zip_counts) <- c("zipcode", "total_calls")

#2 Get data from the Census Bureau showing median household income for Maryland zip codes.
md_median_income_zip <- get_acs(geography = 'zcta', variables = 'B19013_001', state = 'MD', year = 2019)

#3 Join those two dataframes on their zip code columns, starting with the 911 calls dataframe.
combined <- calls_per_zipcode |>
  left_join(md_median_income_zip, by = c("zipcode" = "GEOID"))

scatterplot <- ggplot(data = result_df, aes(x = estimate, y = total_calls, label = zipcode)) +
  geom_point(na.rm = TRUE) +
  geom_text(check_overlap = TRUE, angle = 45, hjust = -.1, size = 3.5) + 
  labs(
    title = "Median Income vs. Total 911 Calls",
    x = "Median Income",
    y = "Total 911 Calls",
    caption = "source: Census Bureau"
  )
scatterplot








# axis on the scatterplot: estimate = x-axis, # of calls = y-axis. Each dot represents a zip code. Use pg county data and md_median_income_zip data. Join two together and create a chart based on that.
```
