---
title: "Reverse Engineering Project"
author: "Morgan Leason"
date: "10/16/23"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are reverse engineering the story, Inside The Political Donation History Of Wealthy Sports Owners (https://fivethirtyeight.com/features/inside-the-political-donation-history-of-wealthy-sports-owners/)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
options(scipen=999)

library(tidyverse)
library(readr)
library(janitor)
library(lubridate)
```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data. 

```{r}
# Load required data

# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")

# Clean required data and prepare for analysis if needed. 

sports_donation <- read_csv("data/sports-political-donations.csv") |> clean_names()

```

## Sentences to Engineer
In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

Who provided the highest number of donations?
A: Charles Johnson. Johnson had 213 total donations.

```{r}
# Put code to reverse engineer sentence here

sports_donation |>
  group_by(owner) |>
  summarise(total_donations = n()) |>
  arrange(desc(total_donations))

# Display results of code below this codeblock

```

### Sentence 2

4. Which league gave the most individual donations? 
A: MLB, which gave 746 donations, well above the rest.

```{r}
# Put code to reverse engineer sentence here
sports_donation |>
  group_by(league) |>
  summarise(total_donation = n()) |>
  arrange(desc(total_donation))
# Display results of code below this codeblock

```

### Sentence 3

5. How many donations for the top party?
A: The top party that received the most donations was the Democratic Party. These results were slightly unclear because the top listing was "Bipartisan," which indicates that some of the donations were tareting something other than the party such as donations for passing a specific bill.

```{r}
# Put code to reverse engineer sentence here
sports_donation |> 
  group_by(party) |>
  summarise(n()) 
# Display results of code below this codeblock

```

### Sentence 4

Which year had the most number of donations?
A: The election year 2018 had the most number of donations.

```{r}
# Put code to reverse engineer sentence here
sports_donation |>
  group_by(election_year) |>
  summarise(count_per_year = n()) |>
  arrange(desc(count_per_year))
# Display results of code below this codeblock

```

### Sentence 5

Which recipient received the largest number of donations?
A: The Gridiron-PAC received the largest number of donations, with a value of 69 donations.

```{r}
# Put code to reverse engineer sentence here
sports_donation |>
  group_by(recipient) |>
  summarise(total_donation = n()) |>
  arrange(desc(total_donation))
# Display results of code below this codeblock

```

-30-