---
title: "Reverse Engineering Project"
author: "Morgan Leason"
date: "10/16/23"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

In this notebook, we are reverse engineering the story, Inside The Political Donation History Of Wealthy Sports Owners (<https://fivethirtyeight.com/features/inside-the-political-donation-history-of-wealthy-sports-owners/>)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
options(scipen=999)

library(tidyverse)
library(readr)
library(janitor)
library(lubridate)
```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data.

```{r}
# Load required data

# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")

# Clean required data and prepare for analysis if needed.

sports_donation <- read_csv("data/sports-political-donations.csv") |>
  mutate(Amount = str_replace_all(Amount,'\\$','')) |>
  mutate(Amount = str_replace_all(Amount, '\\,','')) |>
  mutate(Amount = as.numeric(Amount))

```

## Sentences to Engineer

In this notebook, we are reverse engineering five sentences from the story.

### 1: "Of the 160 owners and commissioners in ESPN's database, 74 contributed a majority of their publicly searchable funds to Republican campaigns, with 48 donating a majority to Democrats and the rest sending most of their money to causes that aren't fully aligned with either party."

```{r}
#number of owners/commissioners —— 158 rows, have to figure out way to get both owners and commisioners for the "more than 160 owners and commisioners"
sports_donation |>
  filter(str_detect(Party, 'Republican')) |>
  group_by(Owner) |> 
  summarise(total = n())
```

### 2: "American professional sports owners have contributed nearly \$47 million in federal elections since 2015, according to research by ESPN in partnership with FiveThirtyEight, including \$10 million to Republican causes and \$1.9 million to Democratic causes so far in the 2020 election cycle."

```{r}
#sum total donations: $46,978,697 ~ nearly $47 million.

sports_donation |>
  summarise(Amount)

#find total Democrat donations specific to 2020 election cycle: $1,877,133 ~ $1.9 million.

sports_donation |>
  filter(str_detect(Party, "Democrat")) |>
  filter(`Election Year`== "2020") |>
  summarise(Amount) |>
  sum()

#find total Republican donations specific to 2020 election cycle: $10,111,284 ~ $10 million.

sports_donation |>
  filter(str_detect(Party, "Republican")) |>
  filter(`Election Year` == "2020") |>
  summarise(Amount) |>
  sum()

```

### 3: "Expanding to all the leagues in our research, here are their totals to each party over the six-year period."

### Which team owners gave the most? Donations from team owners, by cyle and party donated to: (this was in a table in the original story)

```{r}
# Who provided the highest number of donations? Charles Johnson.
sports_donation |>
  group_by(Owner) |>
  summarise(total_donations = n()) |>
  arrange(desc(total_donations))

# Which league gave the most individual donations? MLB
sports_donation |>
  group_by(League) |>
  summarise(total_donation = n()) |>
  arrange(desc(total_donation))

```

### 4. Overall, 51.7 percent of all contributions by WNBA owners went to Democrats, with 42.3 percent going to Republicans. And of the \$1.3 million contributed by WNBA owners to Republicans, 65.5 percent comes from Kelly Loeffler, the Republican from Georgia who is co-owner of the Atlanta Dream.

\`\`\`{r}

```{r}
sports_donation |>
  filter(League == 'WNBA') |>
  group_by(Owner) |>
  summarise(WNBA_AMT = sum(Amount)) |>
  arrange(desc(WNBA_AMT))

#Kelly Loeffler: $896,150

sports_donation |>
  filter(League == 'WNBA') |>
  filter(str_detect(Party, 'Republican')) |>
  summarise(sum(Amount))

#when adding up Republican WNBA spending, the total is $1,055,552 — this is not $1.3m, have to figure out why.

100*(896150/1055552)

sports_donation |>
  filter(League == 'WNBA') |>
  group_by(Party) |>
  summarise(total = sum(Amount))
# there seems to be no WNBA donations for "Bipartisan, but mostly Democrat."

```

### 5. "There are 40 owners who have donated at least \$100,000 to Republican causes over the six-year period, compared to 23 owners topping \$100,000 to Democrats, helping lead to these splits."

```{r}
# There are 37 rows of owners that donated more than $100,000 to Republican causes.
sports_donation |>
  filter(str_detect(Party, 'Republican')) |>
  group_by(Owner) |>
  summarise(total = sum(Amount)) |>
  filter(total >= 100000)

# 22 rows of owners that donated more than $100,000 to Democrat causes.
sports_donation |>
  filter(str_detect(Party, 'Democrat')) |>
  group_by(Owner) |>
  summarise(total = sum(Amount)) |>
  filter(total >= 100000)


# Which year had the most donations? 2018
sports_donation |>
  group_by(`Election Year`) |>
  summarise(count_per_year = n()) |>
  arrange(desc(count_per_year))
```

### 6. "A deep search in the Federal Election Commission database of campaign finances for principal owners, controlling owners, co-owners and commissioners from the NBA, NFL, NHL, WNBA, MLB and NASCAR reveals that this deep-pocketed group has sent \$34.2 million (72.9 percent) to Republican campaigns or super PACs purely supporting Republican\*\* causes, compared to \$10.1 million (21.5 percent) to Democrats over the past three elections. Less than 6 percent of contributions went to bipartisan or unaffiliated recipients."

```{r}
# Money sent to purely Republican causes: 
# sum: 34,246,016 ~ $34.2 million, the total donation amount was $46,978,697
# 34,246,016/46,978,697 = 0.728969 = 72.3%
sports_donation |>
  filter(Party == 'Republican') |>
  summarise(sum(Amount))

100*(34246016/46978697)

# Money sent to purely Democrat causes:
# sum: $10,113,639 ~ $10.1 million.
# $10,113,639/46,978,697 = 0.2152814 = 21.5%

sports_donation |>
  filter(Party == 'Democrat') |>
  summarise(sum(Amount))

100*(10113639/46978697)
```

Data Visualization:

```{r}
sports_donation <- sports_political_donations |>
  group_by(League) |>
  summarise(
    total_amount=n()
  )
```

```{r}
sports_donation |>
  ggplot() +
  geom_bar(aes(x=League, weight=total_amount))
```

```{r}
sports_donation |>
  ggplot() +
  geom_bar(aes(x=League, weight=total_amount)) +
    labs(
    title="Donations by League",
    x = "League",
    y = "Total Amount",
    caption = "source: ESPN FiveThirtyEight"
  ) +
    theme(
    axis.text.x = element_text(angle = 45,  hjust=1)
  )
```
