---
title: "Data Analysis Project"
output: html_notebook
---

```{r}
#Load libraries
library(janitor)
library(tidyverse)
```

```{r}
#Load the data sets and clean the column names
womens_basketball_stats <- read_csv("data/womens_basketball_stats.csv") |> clean_names()

umd_roster <- read_csv("data/umd_roster.csv") |> clean_names()

tabula_womens_offense_defense <- read_csv("data/tabula-womens-offense-defense.csv") |> clean_names()

wbb_majors <- read_csv("data/wbb_majors.csv") |> clean_names()
```

1.  How tall are the players on the University of Maryland women's basketball team? Which player is the tallest?

```{r}
umd_roster |>
  group_by(ht) |>
  summarise(count = n()) |>
  arrange(desc(ht))

```

A: The tallest player is Hawa Doumbouya, with a height of 6 feet, 7 inches.

2.  How do the heights of the players at UMD compare to players outside of UMD - Are UMD centers taller or shorter than the average?

```{r}
#Load and clean the data
iowa_roster <- read_csv('data/Iowa_wbb_roster.csv') |> clean_names() |>
  rename(ht = height)

sc_roster <- read_csv('data/SC_wbb_roster.csv') |> clean_names() |>
  rename(ht = height)

ncstate_roster <- read_csv('data/NCS_wbb_roster.csv') |> clean_names()

ucla_roster <- read_csv('data/UCLA_wbb_roster.csv') |> clean_names()

texas_roster <- read_csv('data/Texas_wbb_roster.csv') |> clean_names()

```

```{r}
#make data sets for each of the five teams
iowa_height <- iowa_roster |>
  group_by(ht) |>
  summarise(count = n()) |>
  arrange(desc(ht))

sc_height <- sc_roster |>
  group_by(ht) |>
  summarise(count = n()) |>
  arrange(desc(ht))

ncs_height <- ncstate_roster |>
  group_by(ht) |>
  summarise(count = n()) |>
  arrange(desc(ht))

ucla_height <- ucla_roster |>
  group_by(ht) |>
  summarise(count = n()) |>
  arrange(desc(ht))

texas_height <- texas_roster |>
  group_by(ht) |>
  summarise(count = n()) |>
  arrange(desc(ht))

```

A: This does show us the heights of everyone on all of these teams each put into different data sets. In these datsets you can clearly see that Maryland have a height problem when it compeares to the top teams in the country. Texas have four players that are 6'4" or taller, and UCLA have six players that are 6'3" or taller. South Carolina have four players 6'3" or taller, NC State also have four, and Iowa have two players 6'3" or taller. All of these team top Maryland's total of one player taller than 6'3".

3.  Which University of Maryland women's basketball players score the highest points in games?

```{r}

womens_basketball_stats |>
  group_by(avg_23, number_player) |>
  summarise(count = n()) |>
  arrange(desc(avg_23))
```

A: Our data shows that Shyanne Sellers averages the most points per game (31.9) followed by Jakia Brown Turner (26.9), Lavender Briggs (23.4), Allie Kubek (22.4) and Brinae Alexander (22.3). To determine this, we sorted the average points per game in the order from most to least, finding these five players at the top as a result.

Join the women's basketball stats and umd roster data sets together:

```{r}
womens_basketball_stats<- womens_basketball_stats |>
  rename(name = "number_player") |>
  separate(name, c('name1', 'name2', 'name3')) |>
  mutate(new_name = paste(name3, name2, sep=" "))

basketball_stats <- umd_roster |>
  inner_join(womens_basketball_stats, by=c("name" = "new_name"))

```

4.  What do the players study at UMD? What major appears to be the most common?

```{r}
womens_basketball_stats <- womens_basketball_stats |>
  inner_join(wbb_majors, by=c("new_name"))

womens_basketball_stats |>
  group_by(new_name, major.x) |>
  summarise(major.x) |>
  arrange(major.x)
```

A: The most common major to study is Letters & Sciences. The findings here are quite clear. The players study majors including Communication, Family Science, Kinesiology, Letters & Sciences and Management studies. Management Studies and Psychology are each tied in second place with two players studying in those fields each. Next steps here could be to look at the data set of the school's majors split up amongst the student population as a whole and see how the team compares.

5. How many women's basketball players are from out of state in comparison to the other schools?

```{r}
# Part one, sort through the data for Maryland.
umd_from_md <- umd_roster |>
  filter(str_detect(hometown_high_school,"MD"))

umd_from_md

#Part Two- A look at South Carolina
sc_from_sc <- sc_roster |>
  filter(str_detect(hometown,"S.C."))

sc_from_sc

#Q5 Part Three- A look at Iowa
iowa_from_iowa <- iowa_roster |>
  filter(str_detect(hometown,"Iowa"))

iowa_from_iowa
```

Answer Q5 Part One: Here we can see that five of the 14 players on the Maryland roster are from the state of Maryland, that's a pretty high percentage. Let's see how it compares to some of the others.

Answer Q5 Part Two: Here we can see that the number one team in the country has only two players on the roster that are from their state. A startlingly low number for such a big and presitigous program in that state.

Answer Q5 Part Three: It's even more surprising that Maryland has more in state recruits than Iowa, a program that's been on the top of the country with Iowan Caitlin Clark for the last few years.

Our most surprising finding so far I believe to be the data when it relates to the amount of in state players that Maryland has. Some would argue that the reason that these top programs aren't filled with in state talent is because they can get better talent from outside soruces. While this is true in some cases, some of these teams have their best pleayer from their state, and still have less in state players than UMD. This shows just how good Maryland is at retaining basketball talent from the DMV area, a place which has a lot of this type of talent.

-   Identify your most newsworthy finding so far and explain why it is.

Our most surprising finding so far I believe to be the data when it relates to the amount of in state players that Maryland has. Some would argue that the reason that these top programs aren't filled with in state talent is because they can get better talent from outside soruces. While this is true in some cases, some of these teams have their best pleayer from their state, and still have less in state players than UMD. This shows just how good Maryland is at retaining basketball talent from the DMV area, a place which has a lot of this type of talent.
